Detection and Classification of Supernova Gravitational Waves Signals: A Deep Learning Approach

### 2019-12-11 main.tex
Added my comments up to the results and discussion section.

\begin{table}[]
\centering
\begin{threeparttable}
\caption{The architecture of the \ac{CNN}}
\label{table:architecture}
\begin{tabular}{ccccc}
\\
\toprule
Layer & Type      & Neurons  & Filter size & Act. Fun \\ \hline
1     & Conv      & 11        & 32          & Elu      \\ 
2     & Max-pool  &          & 8          &          \\
3     & Conv      & 11        & 8          & Elu      \\
4     & Max-pool  &          & 6           &          \\
5     & Conv      & 11       & 6           & Elu      \\
6     & Conv      & 11       & 4           & Elu      \\
7     & Conv      & 13       & 4           & Elu      \\
8     & Conv      & 13       & 4           & Elu      \\
9     & Conv      & 13       & 4           & Elu      \\
10    & Conv      & 13       & 4           & Elu      \\
11    & Max-pool  &          & 2           &          \\
12    & Fully-con & 64(50\%) &             & Elu      \\
13    & Fully-con & 32(50\%) &             & Elu      \\
14    & Fully-con & 3        &             & Softmax      \\ 
\hline
\hline
\end{tabular}
\begin{tablenotes}
\setlength\labelsep{0pt}
\normalfont{
\item The architecture of the \ac{CNN} used in this work for the purpose of
distinguishing supernova signals in backgound \ac{GW} detector noise vs detectr
noise alone. In the table, Conv means convolutional layer, Max-pool means
max-pooling layer, and Fully-con means fully-connected layer. Neuron and Act.
Fun indicate the number of neurons and the activation function used for the
layer respectively. The numbers in the bracket for the fully-connected layers
are the number used for drop-out~\cm{You haven't discussed drop-out yet. You
should mention it earlier when you introduce CNNs}.}
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\begin{table}[]
\centering
\begin{threeparttable}
\caption{Waveforms}
\label{table:waveforms}
\begin{tabular}{rccc}
\toprule
    & Mechanism         & Mass ($\text{M}_\odot$)                         & No.  \\
\hline                         
Abdikamalov\cite{abdikamalov2014measuring} & M   & 12.0                          & 92   \\
Dimmelmeier\cite{dimmelmeier2008gravitational} & M   & 11.2,15.0,20.0,40.0        & 136  \\
Richers\cite{richers2017equation}     & M   & 12.0                          & 1824 \\
Andresen\cite{10.1093mnrasstz990}    & N   & 15.0                          & 6    \\
Kuroda\cite{kuroda2017correlated}      & N   & 11.2,15.0                    & 2    \\
Muller\cite{muller2012parametrized}      & N   & 15.0,20.0                    & 6    \\
Murphy\cite{murphy2009model}      & N   & 12.0,15.0,20.0,40.0        & 16   \\
Ott$_1$\cite{ott2009gravitational}       & N   & 15.0                          & 2    \\
Ott$_2$\cite{ott2013general}       & N   & 27.0                          & 8    \\
Powell\cite{powell2019gravitational}      & N   & 3.5,18.0                      & 2    \\
Radice\cite{radice2019characterizing}      & N   & 9,10,11,12,13,19,25,60 & 8    \\
Yakunin$_1$\cite{yakunin2015gravitational}   & N   & 12,15,20,25                & 4    \\
Yakunin$_2$\cite{yakunin2017gravitational}   & N   & 15                            & 1   \\
\hline
\hline
\end{tabular}
\begin{tablenotes}
\setlength\labelsep{0pt}
\normalfont{
\item The mass ranges and mechanisms of the progenitors of the simulated
waveforms used in this work.  The first column refers to the studies. Mechanism
indicates the explosion mechanism for the waveforms, with M for the
magnetorotational mechanism and N for the neutrino-driven mechanism.  Mass
refers the mass, in units of solar mass, of the progenitors in the simulations.
No. means the number of waveforms available from the study.} 
\end{tablenotes}
\end{threeparttable} 
\end{table}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{mag_waveforms.png}
\caption{Examples of simulated waveforms used in the work for which the
explosion mechanism is magnetorotational mechanism.  The top panel shows a
waveform from~\cite{abdikamalov2014measuring}. The progenitor is
$12\text{M}_\odot$ with differential rotation parameter $A=300\text{km}$. The
initial angular velocity at the core $\Omega_c$ is $8.5\text{rads/s}$. The
middle panel shows a waveform from~\cite{dimmelmeier2008gravitational}. The
simulation corresponds to a progenitor of $15\text{M}_\odot$ with $A=50$km and
$\Omega_c = 1.91\text{rads/s}$.  The lower panel shows a waveform
from~\cite{richers2017equation} where the progenitor is $12\text{M}_\odot$, $A$
and $\Omega_c$ are $10000$km and $0.5\text{rad/s}$ respectively. In all panels,
only the $h_{+}$ polarisations are shown because the simulations are
axis-symmetric or 2D, and therefore described by only one polarisation.  The
sources are assumed to be at a distance of $10$ kpc. The time is not related to
the time at core bounce~\cm{then what is the time in reference to? Also, these
plots are too small and squashed. The text is certainly too small to be
readable and they should all have equal y and x-axis scales for comparison.}.
\label{fig:magwaveforms}} 
\end{figure}

\begin{figure}
\includegraphics[width=0.5\textwidth]{neu_waveforms.png}
\caption{Examples of simulated waveforms used in the work for which the explosion mechanism is neutrino-driven mechanism.
From the top to the bottom, the panels show waveforms from\cite{10.1093mnrasstz990, murphy2009model, radice2019characterizing}, with progenitors of masses equal to $15\text{M}_\odot$,
$15\text{M}_\odot$ and $60\text{M}_\odot$ respectively. 
The sources are assumed to be at $10$ kpc from earth. The time is not related to the time at core bounce.
\label{fig:neuwaveforms}}
\end{figure}
Introducing waveforms from a variety of studies essentially introduces a distribution of the waveforms that covers a larger parameter space, 
making the data samples harder for the \ac{CNN} to learn.
As a result, the performance of the trained \ac{CNN} on the testing samples 
may appear to be worse than that if the data were generated using only a subset of the waveforms.
However, a larger parameter space provides the advantage of forcing the \ac{CNN} to 
learn the features common among the waveforms, rather than a subset of the waveforms.
The trained network can thus perform better when facing waveforms with unexpected features, which is more likely in reality.

Since the waveforms are generated at various distances, sampling rates and durations, 
it is necessary to normalise the waveforms before they can be used for the generation of the time series.
To do this, we first scale the amplitudes of the waveforms by moving the sources to $10$ kpc from earth.
We then ensure that the sampling rate are identical for all the waveforms by down sampling 
them to a pre-selected sampling rate (e.g., $4096$Hz).
The longest duration $\tau$ among the waveforms is then identified and each of the remaining waveforms is padded with zeros
to this duration.
To introduce as less artefact as possible, a high pass filter with a low cut-off frequency equal to $11$Hz 
and a tukey window ($\alpha = 0.08$) are applied prior to the zero padding. 
To balance the difference in the number of waveforms 
between the two mechanisms in the final data set, $36$ copies of each neutrino-driven waveform are duplicated.
After this procedure, the simulated 
waveforms are then $\bold{S}(t) = \{\bold{s}_{1}(t), \bold{s}_{2}(t), ..., \bold{s}_{m}(t)\}$, where $m$ is the number of the waveforms and 
$\bold{s}_{q}$ is the $q^{\text{th}}$ waveform, defined by,
\begin{equation}\label{eq:ht}
 \bold{s}_q(t) = \left(\begin{array}{c}
                        h_q^+(t) \\
                        h_q^\times(t)
                       \end{array}\right),
\end{equation}
where $h_q^+(t)$ and $h_q^\times(t)$ are the two polarisations of the waveform and $t$ the time. 
As mentioned, some of the waveforms used in the work are generated with the simulations being axis-symmetric, the \acp{GW} are entirely described by one polarisation $h_q^+(t)$. 
The corresponding $h_q^\times(t)$ for these waveforms are a zero vector. 
In this work, we perform simulations for distances equal to $10$, $20$, $30$, $40$, $50$, $60$, $80$, $100$, $150$ and $200$ kpc.
If for a training session, the distance $d_\text{L}$ interested is not $10$ kpc, the following relationship is employed to scale the amplitudes of the waveforms,
\begin{equation}\label{eq:sap}
 \bold{s}^{d_\text{L}}_q(t) =  \frac{10\text{kpc}\times\bold{s}_q(t)}{d_\text{L}},
\end{equation}
where $\bold{s}^{d_\text{L}}_q(t)$ is the waveform if the source is at $d_\text{L}$ from earth.

The next step is to generate simulated time series for the \ac{GW} detectors in a network using $\bold{S}(t)$.
%The input data consist of simulated time-series from a network of \ac{GW} detectors.
Since the purpose of building a \ac{CNN} is to categorise an input data into three exclusive classes,
in total three types of time-series are generated.
For the time series containing either a magnetorotational signal or a neutrino-driven signal, 
we start by selecting a waveform $\bold{s}_q$ from $\bold{S}(t)$.
A random location of (right ascension, declination) $=(\alpha, \delta)$ in the sky is selected from a uniform distribution on $\alpha$ and a uniform distribution on the sine of $\delta$.
The corresponding antenna patterns $\bold{F}(\alpha, \delta, t)$, given by the following equation,
\begin{equation}\label{eq:ap}
\bold{F}(\alpha, \delta, t)= \left(\begin{array}{cc}
f_1^+(\alpha, \delta, t)~~f_1^\times(\alpha, \delta, t) \\
\vdots~~~~~~~~~~~~\vdots \\
f_k^+(\alpha, \delta, t)~~f_k^\times(\alpha, \delta, t) \\
\end{array}\right),
\end{equation}
is computed, 
where $f^+(\alpha, \delta, t)$ and $f^{\times}(\alpha, \delta, t)$ are the antenna pattern functions for the two polarizations and $k$ is the number of detectors as defined above.
The relative delays in the arrival times for the location are also computed and applied to the selected waveform.
The delay in arrival time between a detector and the center of the earth is given by,
\begin{equation}\label{eq:tdetector}
%
\Delta t = \frac{\bold{n} \cdot \bold{r}}{c},
%
\end{equation}
%
where $\bold{n}$ is the propagation direction of the \ac{GW}, $c$ the speed of light, and $\bold{r}$ the location vector of the detector
relative to the center of the Earth. 
The resulting signal $ \bold{h}_j(\alpha, \delta, t)$ as received at the detectors is then given by the following equation,
\begin{equation}\label{eq:ht2}
 \bold{h}_j(\alpha, \delta, t) = \bold{F}(\alpha, \delta, t) \times \bold{s}_q(t),
\end{equation}
where the time delays $\Delta t$ are absorbed into the corresponding notations and the subscript $j$ is defined in Eq.\ref{eq:cce}.
Next, we generate independent Gaussian noise $\bold{N}_j(t) = \{\bold{n}_{1j}(t), \bold{n}_{2j}(t), ..., \bold{n}_{kj}(t)\}'$ 
for each detector in the network using their respective power spectral densities.  
In the above notation, the symbol $'$ indicates transpose. 
%It is worth mentioning that although in reality, 
%noise at \ac{GW} detectors may not be entirely Gaussian and contaminated with glitches,
%the application of \ac{CNN} is actually not limited to time series where the noise is Gaussian.
%Although  we will focus on Gaussian noise in this work, 
%a \ac{CNN} can in fact be trained to identify the presence of a glitch in the data \cite{mukund2017transient, zevin2017gravity, george2017deep, gabbard2018matching}.
The duration of the generated noise is $1.6$ times longer than $ \bold{h}_j(\alpha, \delta, t)$. 
A random number $x$ will then be generated determining where in the generated noise the signal will be placed, as given by,
\begin{equation}
\bold{d}_j(\alpha, \delta, t, x)=
\begin{cases}
\bold{N}_j(t)~~~~~~~~~~~~~~~~~t < x; \\
\bold{N}_j(t) + \bold{h}_j(\alpha, \delta, t)~x \leq t \leq \tau + x; \\
\bold{N}_j(t)~~~~~~~~~~~~~~~~~t > \tau + x. \\
\end{cases}
\end{equation}
This is to avoid the possibility that the \ac{CNN} learns human artefact instead of common features of the waveforms by having the signals always starting at the same place.
For each time series of the background noise class, independently simulated Gaussian background noise of the same duration as that of the other classes 
are generated for each detectors in the network.
This means a data sample for this class is defined as,
\begin{equation}\label{eq:noise}
 \bold{d}_j(t) = \bold{N}_j(t).
\end{equation}

Since training a \ac{CNN} of a given structure to its maximum capacity requires many data samples, we augment our data set by iterating over $\bold{S}(t)$ until 
for each waveform in $\bold{S}(t)$, we have $31$ data samples $\bold{d}$ generated using the procedure described above.
This step is essential for the \ac{CNN} to learn the features of the waveforms and to identify them effectively under different noise scenarios.
After that, we generate independent noise realisations for the background noise class.
The entire data set $\bold{D}$ defined below, has roughly $l = 1.8 \times 10^{5}$ data samples, where each class has approximately $6\times10^{4}$ data samples.
\begin{equation}\label{eq:finaldata}
\bold{D}(\boldsymbol{\alpha}, \boldsymbol{\delta}, \boldsymbol{\delta_t}, \bold{t}, \bold{x}) =  \left(\begin{array}{c}
\bold{d}_1(\alpha, \delta, \delta_t, t, x) \\
\vdots \\
\bold{d}_o(\alpha, \delta, \delta_t, t, x) \\
\vdots \\
\bold{d}_l(t) \\
\end{array}\right).
\end{equation}
In the above equation, the subscript $o$ is the sum of the data samples in which there is a \ac{GW} signal in the time series. 
The final step is to whiten $\bold{D}$ using the power spectral densities.
In Figure \ref{fig:sample}, a representative example  of time series for one detector is shown. 
\begin{figure}
\includegraphics[width=0.5\textwidth]{datasample.png}
\caption{Representative example of the simulated time series used to train/validate/test the \ac{CNN}.
The blue shows a whitened time series with a signal buried in Gaussian noise of unit variance.
The red shows the same signal free of noise and whitened. 
\label{fig:sample}}
\end{figure}
A data sample consists of such time series of the same signal or background noise from all the detectors in the network.
The data samples are then split, with $1\times10^{4}$ samples being randomly taken for validation, $10\%$ of the rest for testing and $90\%$ for training.
When the training is finished for a distance, 
the above described procedure will be repeated for another distance until the training
for all distances have been carried out for a network.
The entire procedure is then repeated for another network of \ac{GW} detectors.
As mentioned previously, the \ac{CNN} will be trained for two networks of \ac{GW} detectors.
We present the networks in Table \ref{table:network}. For the remaining of the paper, we will use their acronyms to refer to the networks.
\begin{table}[]
\centering
\begin{threeparttable}
\caption{Detector networks}
\label{table:network}
\begin{tabular}{ccc}
\toprule
Network            & Detectors                & Acronym                 \\
\hline
\multirow{4}{*}{1} & \ac{aLIGO} H*    & \multirow{4}{*}{HLVK}   \\
                   & \ac{aLIGO} L* &                         \\
                   & \ac{AdVirgo}           &                         \\
                   & KAGRA                    &                         \\
                   \hline
\multirow{4}{*}{2} & LIGO A+ H          & \multirow{4}{*}{H+L+VK} \\
                   & LIGO A+ L       &                         \\
                   & \ac{AdVirgo}           &                         \\
                   & KAGRA                    &                         \\
\hline
\hline
\end{tabular}
\begin{tablenotes}
\setlength\labelsep{0pt}
\normalfont{
\item The networks of detectors used in this work.\\
* H refers to the detector in Hanford and L the detector in Livingston.}
\end{tablenotes}
\end{threeparttable}
\end{table}

>>>>

%
% table of CNN hyper-parameters
%
\begin{table}[]
\centering
\begin{threeparttable}
\caption{The architecture of the \ac{CNN}}
\label{table:architecture}
\begin{tabular}{ccccc}
\\
\toprule
Layer & Type      & Neurons  & Filter size & Act. Fun \\ \hline
1     & Conv      & 11        & 32          & Elu      \\ 
2     & Max-pool  &          & 8          &          \\
3     & Conv      & 11        & 8          & Elu      \\
4     & Max-pool  &          & 6           &          \\
5     & Conv      & 11       & 6           & Elu      \\
6     & Conv      & 11       & 4           & Elu      \\
7     & Conv      & 13       & 4           & Elu      \\
8     & Conv      & 13       & 4           & Elu      \\
9     & Conv      & 13       & 4           & Elu      \\
10    & Conv      & 13       & 4           & Elu      \\
11    & Max-pool  &          & 2           &          \\
12    & Fully-con & 64(50\%) &             & Elu      \\
13    & Fully-con & 32(50\%) &             & Elu      \\
14    & Fully-con & 3        &             & Softmax      \\ 
\hline
\hline
\end{tabular}
\begin{tablenotes}
\setlength\labelsep{0pt}
\normalfont{
\item The architecture of the \ac{CNN} used in this work for the purpose of
distinguishing supernova signals in backgound \ac{GW} detector noise vs detectr
noise alone. In the table, Conv means convolutional layer, Max-pool means
max-pooling layer, and Fully-con means fully-connected layer. Neuron and Act.
Fun indicate the number of neurons and the activation function used for the
layer respectively. The numbers in the bracket for the fully-connected layers
are the number used for drop-out~\cm{You haven't discussed drop-out yet. You
should mention it earlier when you introduce CNNs}.}
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

%
% table of simulation paramaters
%
\begin{table}[]
\centering
\begin{threeparttable}
\caption{Waveforms}
\label{table:waveforms}
\begin{tabular}{rccc}
\toprule
    & Mechanism         & Mass ($\text{M}_\odot$)                         & No.  \\
\hline                         
Abdikamalov\cite{abdikamalov2014measuring} & M   & 12.0                          & 92   \\
Dimmelmeier\cite{dimmelmeier2008gravitational} & M   & 11.2,15.0,20.0,40.0        & 136  \\
Richers\cite{richers2017equation}     & M   & 12.0                          & 1824 \\
Andresen\cite{10.1093mnrasstz990}    & N   & 15.0                          & 6    \\
Kuroda\cite{kuroda2017correlated}      & N   & 11.2,15.0                    & 2    \\
Muller\cite{muller2012parametrized}      & N   & 15.0,20.0                    & 6    \\
Murphy\cite{murphy2009model}      & N   & 12.0,15.0,20.0,40.0        & 16   \\
Ott$_1$\cite{ott2009gravitational}       & N   & 15.0                          & 2    \\
Ott$_2$\cite{ott2013general}       & N   & 27.0                          & 8    \\
Powell\cite{powell2019gravitational}      & N   & 3.5,18.0                      & 2    \\
Radice\cite{radice2019characterizing}      & N   & 9,10,11,12,13,19,25,60 & 8    \\
Yakunin$_1$\cite{yakunin2015gravitational}   & N   & 12,15,20,25                & 4    \\
Yakunin$_2$\cite{yakunin2017gravitational}   & N   & 15                            & 1   \\
\hline
\hline
\end{tabular}
\begin{tablenotes}
\setlength\labelsep{0pt}
\normalfont{
\item The mass ranges and mechanisms of the progenitors of the simulated
waveforms used in this work.  The first column refers to the studies. Mechanism
indicates the explosion mechanism for the waveforms, with M for the
magnetorotational mechanism and N for the neutrino-driven mechanism.  Mass
refers the mass, in units of solar mass, of the progenitors in the simulations.
No. means the number of waveforms available from the study.} 
\end{tablenotes}
\end{threeparttable} 
\end{table}

@@

%
% plot of example magneto waveforms
%
\begin{figure}
\includegraphics[width=0.5\textwidth]{mag_waveforms.png}
\caption{Examples of simulated waveforms for which the explosion is modeled by
the magnetorotational mechanism. The top panel shows a waveform
from~\cite{abdikamalov2014measuring}. The progenitor is $12\text{M}_\odot$ with
differential rotation parameter $A=300\text{km}$. The initial angular velocity
at the core $\Omega_c$ is $8.5\text{rads/s}$. The middle panel shows a waveform
from~\cite{dimmelmeier2008gravitational}. The simulation corresponds to a
progenitor of $15\text{M}_\odot$ with $A=50$km and $\Omega_c =
1.91\text{rads/s}$.  The lower panel shows a waveform
from~\cite{richers2017equation} where the progenitor is $12\text{M}_\odot$, $A$
and $\Omega_c$ are $10000$km and $0.5\text{rad/s}$ respectively. In all panels,
only the $h_{+}$ polarisations are shown because the simulations are
axis-symmetric or 2D, and therefore described by only one polarisation.  The
sources are assumed to be at a distance of $10$ kpc. The time is not related to
the time at core bounce~\cm{then what is the time in reference to? Also, these
plots are too small and squashed. The text is certainly too small to be
readable and they should all have equal y and x-axis scales for comparison.}.
\label{fig:magwaveforms}} 
\end{figure}

%
% plot of example neutrino waveforms
%
\begin{figure}
\includegraphics[width=0.5\textwidth]{neu_waveforms.png}
\caption{Examples of simulated waveforms for which the explosion mechanism is
modeled by the neutrino-driven mechanism. From the top to the bottom, the
panels show waveforms from~\cite{10.1093mnrasstz990, murphy2009model,
radice2019characterizing}, with progenitors of masses equal to
$15\text{M}_\odot$, $15\text{M}_\odot$ and $60\text{M}_\odot$ respectively.
The sources are assumed to be at $10$ kpc from earth. The time is not related
to the time at core bounce.~\cm{then what is the time in reference to? Also, these
plots are too small and squashed. The text is certainly too small to be
readable and they should all have equal y and x-axis scales for comparison. Why
does the x-axis extend to longer than the longest waveform?}
\label{fig:neuwaveforms}}
\end{figure}

%
% discuss the reason of using multiple waveform families
%
Introducing waveforms from a variety of studies introduces a distribution of
the waveforms that covers a larger parameter space, making the data samples
harder for the \ac{CNN} to learn. As a result, the performance of the trained
\ac{CNN} on the testing samples may appear to be worse than that if the data
were generated using only a subset of the waveforms. However, a larger
parameter space provides the advantage of forcing the \ac{CNN} to learn the
features common among the waveforms rather than simply a subset of the
waveforms.  The trained network can thus perform better when presented with
waveforms with unexpected features, which is more likely in reality.

%
% describe data pre-processing
%
Since the waveforms are generated at various distances, sampling rates and
durations, it is necessary to normalise the waveforms before they can be used
for the generation of the time-series. To do this, we first scale the
amplitudes of the waveforms by moving the sources to $10$ kpc from earth. We
then ensure that the sampling rate are identical for all the waveforms by down
sampling them to a pre-selected sampling rate (e.g., $4096$Hz~\cm{e.g. means
for example, is this an example rate or is it THE rate?}). The longest duration
$\tau$ among the waveforms is then identified and each of the remaining
waveforms is padded with zeros to this duration. To introduce as few artefacts
as possible, a high pass filter with a low cut-off frequency equal to $11$Hz
and a tukey window ($\alpha = 0.08$) are applied prior to the zero padding.

%
% data augmentation
%
To balance the difference in the number of waveforms between the two mechanisms
in the final data set~\cm{what is the final dataset? Is it the training,
validation or test, or all of them? Please be clearer}, $36$ copies of each neutrino-driven waveform are
made~\cm{please explain this in more detail, why is this done and how is this
done. Surely they are not simply exact copies? Does each copy get a different
noise realisation or different arrival time? This leaves many questions. Do
you simply make the number of magneto and neutrino wavelforms equal? What about
the sub categories? Do you still have 8 times as many Ott2 waveforms as
Yakunin2? If so, why?}. After this procedure, the
simulated waveforms are then $\bold{S}(t) = \{\bold{s}_{1}(t), \bold{s}_{2}(t),
..., \bold{s}_{m}(t)\}$, where $m$ is the number of the waveforms and
$\bold{s}_{q}$ is the $q^{\text{th}}$ waveform, defined by, 
%
\begin{equation}\label{eq:ht}
\bold{s}_q(t) = \left(\begin{array}{c} h_q^+(t) \\ h_q^\times(t)
\end{array}\right), 
\end{equation} 
%
where $h_q^+(t)$ and $h_q^\times(t)$ are the two polarisations of the waveform
and $t$ is the time. As mentioned, some of the waveforms used in the work are
generated with the simulations being axisymmetric, in which case the waveforms
are entirely described by one polarisation $h_q^+(t)$. The corresponding
$h_q^\times(t)$ for these waveforms are defined as a vector of zeros.~\cm{This
slightly worries me that classification between models may be artifically
enhanced by forcing the cross polarisation to be zero. The network might be
able to easily use this to distinguich between classes when in reality
waveforms will not show this property. A referee may have the sam concern.}

%
% rescaling the amplitudes for differnet distances
%
In this work, we perform simulations for distances equal to $10$, $20$, $30$,
$40$, $50$, $60$, $80$, $100$, $150$ and $200$ kpc. If for a training
procedure, the default distance $d_\text{L}$ of a waveform is not $10$ kpc, the
following relationship is employed to scale the amplitudes of the waveforms,

\begin{equation}\label{eq:sap} 
\bold{s}^{d_\text{L}}_q(t) =
\frac{10\text{kpc}\times\bold{s}_q(t)}{d_\text{L}}, 
\end{equation} 
%
where $\bold{s}^{d_\text{L}}_q(t)$ is the waveform for the source at $d_\text{L}$
from earth.~\cm{this equation and preceeding text is kind of obvious and could
be said in far less words. You just simply rescale the amplitude by the
distance since the amplitude is inversely proportional to the distance.}

%
% Apply antenna patterns
%
The next step is to generate simulated time-series for the \ac{GW} detectors in
a network using $\bold{S}(t)$.
%The input data consist of simulated time-series from a network of \ac{GW}
%detectors.
Since the purpose of building a \ac{CNN} is to categorise an input data sample
into three exclusive classes, in total three types of time-series are
generated.  For the time series containing either a magnetorotational signal or
a neutrino-driven signal, we start by selecting a waveform $\bold{s}_q$ from
$\bold{S}(t)$. A random location of (right ascension, declination) $=(\alpha,
\delta)$ in the sky is selected from a uniform distribution on $\alpha$ and a
uniform distribution on $\sin\delta$. The corresponding antenna patterns
$\bold{F}(\alpha, \delta, t)$, given by the following equation~\cm{the
following equation doesn't give you the antenna patterns. It just lists them as
an array. Do you really need this array format to explain the waveforms? Does
it help you?},
%
\begin{equation}\label{eq:ap}
\bold{F}(\alpha, \delta, t)= \left(\begin{array}{cc}
f_1^+(\alpha, \delta, t)~~f_1^\times(\alpha, \delta, t) \\
\vdots~~~~~~~~~~~~\vdots \\
f_k^+(\alpha, \delta, t)~~f_k^\times(\alpha, \delta, t) \\
\end{array}\right),
\end{equation}
%
is computed, where $f^+(\alpha, \delta, t)$ and $f^{\times}(\alpha, \delta, t)$
are the antenna pattern functions for the two polarizations and $k$ is the
number of detectors as defined above~\cm{How do you select $t$ and what does it
represent?}.

%
% the time delays
%
The relative delays in the arrival times of the signals at each
detector are a function of the sky location and are also computed and applied
to the selected waveform. The delay in arrival time between a detector and the
center of the earth is given by,
%
\begin{equation}\label{eq:tdetector}
\Delta t = \frac{\bold{n}(\alpha,\delta) \cdot \bold{r}}{c},
\end{equation}
%
where $\bold{n}(\alpha,\delta)$ is the propagation direction of the \ac{GW},
$c$ is the speed of light, and $\bold{r}$ the location vector of the detector
relative to the center of the Earth. The resulting $j$'th signal $\bold{h}_j(\alpha,
\delta, t)$ received by the detector network is then described by,
%
\begin{equation}\label{eq:ht2}
 \bold{h}_j(\alpha, \delta, t) = \bold{F}(\alpha, \delta, t) \times
\bold{s}_q(t+\bold{\Delta t}),
\end{equation}
%
where the time delays $\Delta t$ are absorbed into the corresponding notations
and the subscript $j$ is defined in Eq.\ref{eq:cce}~\cm{I changed the equation
to be more explicit so you don't have to say this previous sentence.}.

%
% adding noise
%
Next, we generate independent Gaussian noise $\bold{N}_j(t) =
\{\bold{n}_{1j}(t), \bold{n}_{2j}(t), ..., \bold{n}_{kj}(t)\}'$ for each
detector in the network using their respective power spectral
densities~\cm{give references and name the detector epoch, e.g., advanced
design, O2, 3rd gen, etc..}. In the above notation, the symbol $'$ indicates
transpose~\cm{why is the noise transpose and not the signal? Maybe I'll find
out as I keep reading.}. 
%It is worth mentioning that although in reality, noise at \ac{GW} detectors
%may not be entirely Gaussian and contaminated with glitches, the application
%of \ac{CNN} is actually not limited to time series where the noise is
%Gaussian.  Although  we will focus on Gaussian noise in this work, a \ac{CNN}
%can in fact be trained to identify the presence of a glitch in the data
%\cite{mukund2017transient, zevin2017gravity, george2017deep,
%gabbard2018matching}.
The duration of the generated noise is $1.6$ times longer than
$\bold{h}_j(\alpha, \delta, t)$~\cm{Don't start off explanations like this. It
makes the reader think that you are simply declaring something as fact without
explaining it}. A random number $x$~\cm{what kind of random number? I assume
uniform but between what limits?} will then be generated determining where in
the generated noise the signal will be placed, as given by,
%
\begin{equation}
\bold{d}_j(\alpha, \delta, t, x)=
\begin{cases}
\bold{N}_j(t)~~~~~~~~~~~~~~~~~t < x; \\
\bold{N}_j(t) + \bold{h}_j(\alpha, \delta, t)~x \leq t \leq \tau + x; \\
\bold{N}_j(t)~~~~~~~~~~~~~~~~~t > \tau + x. \\
\end{cases}
\end{equation}
%
~\cm{please define $\tau$.} This is to avoid the possibility that the \ac{CNN}
learns the human artefact instead of common features of the waveforms by having
the signals always starting at the same place.~\cm{well kind of. The reason you
randomise is that the start time is an unknown parameter so you simply want the
network to learn to generalise to deal with that kind of variation. Fixing the
start time at the stame value each time is indeed a human feature but so is
fixing the distance and all the other things that you fix.} For each
time-series of the background noise class, independently simulated Gaussian
background noise~\cm{you need to make it clear that you are also generating
this noise from the detector PSDs. You could also just say that the background
noise class is produced identically but with a zero amplitude signal.} of the
same duration as that of the other classes are generated for each detector in
the network. This means a data sample for this class is defined as,
%
\begin{equation}\label{eq:noise}
 \bold{d}_j(t) = \bold{N}_j(t).
\end{equation}

%
% data augmentation
%
As a general rule, the performance of a \ac{CNN} can always be enhanced by
increasing the volume of training data. Therefore we augment our data set by
iterating over $\bold{S}(t)$ until for each waveform in $\bold{S}(t)$, we have
$31$ data samples $\bold{d}$ generated using the procedure described above.
This step is essential for the \ac{CNN} to learn the features of the waveforms
and to identify them effectively under different noise scenarios and starting
times.  After that, we generate independent noise realisations for the
background noise class. The entire data set $\bold{D}$ defined below, has
roughly $l = 1.8 \times 10^{5}$ data samples, where each of the 3 classes
contains approximately $6\times10^{4}$ data samples.
%
\begin{equation}\label{eq:finaldata}
\bold{D}(\boldsymbol{\alpha}, \boldsymbol{\delta}, \boldsymbol{\delta_t}, \bold{t}, \bold{x}) =  \left(\begin{array}{c}
\bold{d}_1(\alpha, \delta, \delta_t, t, x) \\
\vdots \\
\bold{d}_o(\alpha, \delta, \delta_t, t, x) \\
\vdots \\
\bold{d}_l(t) \\
\end{array}\right).
\end{equation}
%
In the above equation, the subscript $o$ is the sum of the data samples in
which there is a \ac{GW} signal in the time series.~\cm{this notation is very
confusing and not really required. Why do you have to stack the waveforms up
like this? Having to define things like $o$ and $l$ seems messy.} The final
step is to whiten $\bold{D}$ using the power spectral densities~\cm{If you are
going to spend effort explaining how you rescale distances (a very simple
procedure) then you should definitley describe how you whiten the data}.

%
% the general training procedure
%
In Fig.~\ref{fig:sample}, we show a representative example of a pre-processed
input time-series for a single detector. A data sample consists of such a
time-series containing a signal from the same source or simply background noise
from all the detectors in the network. The data samples are then split into 3
groups, with $1\times10^{4}$ samples being randomly selected for validation,
$10\%$ of the remainder for testing and $90\%$ for training. When the training
is finished for a dataset generated at a given distance, the above described
procedure will be repeated for another distance until the training for all
distances have been carried out for a network~\cm{You need to make it clear
that you are training different networks (with the same architecture) for each
distance and not iterativley training the same single network.}. The entire
procedure is then repeated for another network~\cm{the word network is being
used to describe the CNN and the IFOs, it might be a bit confusing} of \ac{GW}
detectors. As mentioned previously~\cm{when was it mentioned previoulsy? give a
ref to the section.}, the \ac{CNN} will be trained for the two networks of
\ac{GW} detectors presented in Table~\ref{table:network}. For the remaining of
the paper, we will use their acronyms to refer to the networks~\cm{what are
these acronymns, state them here.}.

%
% plot of representative data sample
%
\begin{figure}
\includegraphics[width=0.5\textwidth]{datasample.png}
\caption{A representative example of the simulated time-series used to
train/validate/test the \ac{CNN}. The blue shows a whitened time-series with a
signal added to Gaussian noise of unit variance~\cm{if you've whitened it then
of course the noise is zero mean and unit variance - that's the definition so
no need to say it twice.}. The red curve shows the same whitened signal free of
noise.~\cm{what waveform is this and what is it's SNR? Please state this in the
caption. Is this a representative SNR that you can detect? If not then why show
it.}\label{fig:sample}} 
\end{figure}

\begin{table}[]
\centering
\begin{threeparttable}
\caption{Detector networks}
\label{table:network}
\begin{tabular}{ccc}
\toprule
Network            & Detectors                & Acronym                 \\
\hline
\multirow{4}{*}{1} & \ac{aLIGO} H*    & \multirow{4}{*}{HLVK}   \\
                   & \ac{aLIGO} L* &                         \\
                   & \ac{AdVirgo}           &                         \\
                   & KAGRA                    &                         \\
                   \hline
\multirow{4}{*}{2} & LIGO A+ H          & \multirow{4}{*}{H+L+VK} \\
                   & LIGO A+ L       &                         \\
                   & \ac{AdVirgo}           &                         \\
                   & KAGRA                    &                         \\
\hline
\hline
\end{tabular}
\begin{tablenotes}
\setlength\labelsep{0pt}
\normalfont{
\item The networks of detectors used in this work.\\
* H refers to the detector in Hanford and L the detector in Livingston.~\cm{you
* have enough space to simply use Hanford and Livingston in the table itslef.
* You could also have a column for references to the detector papers.}}
\end{tablenotes}
\end{threeparttable}
\end{table}
